---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
As the file is zipped, we will use the function as argument into read.table
```{r}
data <- read.csv(unz("activity.zip", "activity.csv"), header=T)
head(data)
```
Converting Date strings into R date format
```{r}
data$date <- as.Date(data$date,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?
Let´s remove the NA values for this analysis
```{r}
sdata<-data[!is.na(data$steps),]
```
Calculating number of steps for each day
```{r}
steps_per_day<-tapply(sdata$steps, sdata$date, sum)
head(steps_per_day)

```
Creating the histogram
```{r}
hist(steps_per_day,breaks = 10, xlab = "Steps per Day", 
     main = "Frequency of Number of Steps per day")
```


Sumarizing the mean and meadian total steps per day
```{r}
summary(steps_per_day)
```
## What is the average daily activity pattern?
Let´s calculate the average steps by each interval
```{r}
steps_per_interval<-tapply(sdata$steps, sdata$interval, mean)
head(steps_per_interval)

```
Making a timeseries plot
```{r}
xnames <- names(steps_per_interval)
plot(steps_per_interval, type='l', xlab="5-min Interval",xaxt='n',
     ylab = "Average Steps",main = "Average steps by 5 minues interval")
axis(1,at = seq(1,288,by=48),labels=xnames[seq(1,288,by=48)])

```

The time on which maximun number of steps occurs can be found by
```{r}
xnames[which.max(steps_per_interval)]
```

## Imputing missing values
Let´s Calculate and report the total number of missing values in the dataset 
```{r}
sum(is.na(data))
```
Create a new dataset with missing values replaced by the mean of that interval
```{r}
index<-is.na(data)
data2<-data
data2[index,"steps"]<-steps_per_interval[match(data[index,"interval"],xnames)]
```
Calculate the new steps by day with replaced missing values and build the histogram
```{r}
steps_per_day2<-tapply(data2$steps, data2$date, sum)
hist(steps_per_day2,breaks = 10, xlab = "Steps per Day", 
     main = "Frequency of Number of Steps per day")
```

The histogram looks like excalty the same shape as before,but with higher frequencies.
Let's summarize both data with and without missing values to see  the differences
```{r}
summary(steps_per_day)
summary(steps_per_day2)
## Show in a Table
data.frame(Mean = rbind(mean(steps_per_day),mean(steps_per_day2)),
               Median = rbind(median(steps_per_day),median(steps_per_day2)),
               row.names = c("w/NA","w/o NA"))

```
We can see that there is no significant difference if we include the mean values of the time interval as missing data. As expected the mean is still de same, and a slightly change on Median Values 

## Are there differences in activity patterns between weekdays and weekends?

Let´s create a factor with two levels, for weekdays and weekdends
```{r}
weekends <- c("Saturday","Sunday")
data2$Weekdays<-factor(weekdays(data2$date) %in% weekends,
                       labels = c("Weekend","weekday"),
                       levels = c("TRUE","FALSE"))
summary(data2$Weekdays)                       
```
Let´s aggregate data into weekdays and time intervals.
```{r}
data3<-aggregate(list(Mean = data2$steps),
                 by=list(interval=data2$interval,week=data2$Weekdays),
                 FUN = mean)
```

Let´s make a plot using ggplot2
```{r}
library(ggplot2)
ggplot(data3,aes(interval,Mean))+facet_wrap(week~.,nrow=2 )+geom_line()+
labs(y = "Average Steps")

        
```
